---
title: "Investigating the correlations of phenotyper score with clinical data in the NACC data"
output: html_notebook
---

This will investigate the relationship between the mean phenotype scores and clinical scores in the NACC dataset.

```{r}
library(dplyr)
library(stringr)
library(ggplot2)
library(DT)
library(brms)
library(bayesplot)
library(tidybayes)
library(xtable)
library(janitor)
library(tidyr)
library(rstan)
library(ggplot2)
library(xtable)

options(mc.cores = parallel::detectCores())
set.seed(5)
```

```{r}
# import diagnoses
clinical_data_file = "../data/nacc_collatedResults.csv"
clindata = read.table(clinical_data_file, header = TRUE)

clindata_summary <- clindata %>%
  group_by(diagnosis) %>%
  #summarise(n = length(subj)) %>%
  summarise(n = length(subj),
            age = paste(round(mean(age, na.rm = TRUE), digits = 1),
                              paste("(", round(sd(age, na.rm = TRUE), digits = 1), ")", sep = "")),
            sex = paste(length(gender[gender=="Male"]),
                           length(gender[gender=="Female"]),
                           sep = "/")) %>%
  adorn_totals("row")

datatable(clindata_summary)

clindata_summary_xtable <- xtable(clindata_summary)
print(clindata_summary_xtable)
```


```{r}
# remove volume and thickness measures
clindata <- clindata %>%
  dplyr::select(colnames(.)[!grepl("volume", names(.))]) %>%
  dplyr::select(colnames(.)[!grepl("thickness", names(.))])

# import results
# the '50' is the dropout results
results_file = "../results/latest_output_nacc_50.csv"
results = read.table(results_file, header = TRUE, sep = ",")
results <- results %>%
  mutate(subj = paste("sub-NACC", str_pad(nacc_id, width=6, side="left", pad="0"), sep = "")) %>%
  dplyr::select(-nacc_id)

# diagnosis groups

diagnosis_map = list('Alzheimer’s disease (AD)' = 'AD',
                     'Vascular brain injury or vascular dementia including stroke' = 'OD',
                     'Lewy body disease (LBD)' = 'OD', 
                     'Control' = 'Control', 
                     'Depression' = 'OND',
                     'Missing/unknown' = 'remove', ## <<---------- !
                     'Other neurologic, genetic, or infectious condition' = 'OND',
                     'Cognitive impairment for other specified reasons (i.e., written-in values)' = 'OND',
                     'Prion disease (CJD, other)' = 'OD', 
                     'FTLD, other' = 'OD', 
                     'Anxiety disorder' = 'OND',
                     'Corticobasal degeneration (CBD)' = 'OD',
                     'Cognitive impairment due to medications' = 'OND',
                     'Other psychiatric disease' = 'OND',
                     'Progressive supranuclear palsy (PSP)' = 'OD',
                     'Cognitive impairment due to systemic disease or medical illness' = 'OND',
                     'Traumatic brain injury (TBI)' = 'OND',
                     'Cognitive impairment due to alcohol abuse' = 'OND', 
                     'Bipolar disorder' = 'OND',
                     'Schizophrenia or other psychosis' = 'OND',
                     'FTLD with motor neuron disease (e.g., ALS)' = 'OD')

results <- results %>%
  left_join(clindata, by = "subj") %>%
  mutate(diagnosis = diagnosis_map[diagnosis]) %>%
  filter(diagnosis != "remove")

results$diagnosis <- factor(unlist(results$diagnosis),
                            levels = c("Control", "AD", "OD", "OND"))

rm(clindata)

results_summary <- results %>%
  group_by(diagnosis) %>%
  summarise(n = length(subj),
            age = paste(round(mean(age, na.rm = TRUE), digits = 1),
                              paste("(", round(sd(age, na.rm = TRUE), digits = 1), ")", sep = "")),
            sex = paste(length(gender[gender=="Male"]),
                           length(gender[gender=="Female"]),
                           sep = "/")) %>%
  adorn_totals("row")

datatable(results_summary)

results_summary_xtable <- xtable(results_summary)
print(results_summary_xtable, include.rownames = FALSE)
```

```{r}
# add column for AD-like positive/negative
results <- results %>%
  group_by(subj) %>%
  mutate(adlike = if (mean > 0.5) {"Positive"} else {"Negative"})

# set graph limits for breakpoint analysis plots
bp_y_min = -1.2
bp_y_max = 0.7
```

# MMSE

First we'll plot MMSE against the mean

```{r}
clinscore_name = "MMSE"

naccmmse <- results %>%
  filter(!is.na(nacmmse))

naccmmse_summary <- naccmmse %>%
  group_by(adlike) %>%
  summarise(n = length(subj),
            mmse = paste(round(mean(nacmmse), digits = 1),
                         " (", round(sd(nacmmse), digits = 1), ")",
                         sep = ""),
            age = round(mean(age, na.rm = TRUE), digits = 1),
            sex = paste(length(gender[gender == "Female"]),
                        length(gender[gender == "Male"]),
                        sep = "/"))

datatable(naccmmse_summary)
```

```{r}
p <- ggplot(naccmmse,
            aes(y = nacmmse,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
p_clinscore <- ggplot(naccmmse, aes(x = mean, y = nacmmse, group = adlike, colour = adlike))
p_clinscore <- p_clinscore + geom_point() + geom_smooth(method = lm) +
  labs(x = "AD score",
       y = clinscore_name) +
  theme(legend.position = "None")
print(p_clinscore)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(nacmmse ~ adlike + age, 
                data = naccmmse, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)


```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(naccmmse[naccmmse$adlike == "Negative", "nacmmse"][["nacmmse"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

```{r}
# z-score values
mean_age = mean(naccmmse$age)
sd_age = sd(naccmmse$age)
naccmmse$age <- (naccmmse$age - mean_age) / sd_age

mean_nacmmse = mean(naccmmse$nacmmse)
sd_nacmmse = sd(naccmmse$nacmmse)
naccmmse$nacmmse <- (naccmmse$nacmmse - mean_nacmmse) / sd_nacmmse

mean_mean = mean(naccmmse$mean)
sd_mean = sd(naccmmse$mean)
naccmmse$mean <- (naccmmse$mean - mean_mean) / sd_mean

```

```{r}

# set priors
mmse_priors = c(set_prior("cauchy(0,1)", class = "b", coef = "age"),
                set_prior("cauchy(0,1)", class = "b", coef = "adlikePositive"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean:adlikePositive"))

fit_mmse <- brm(nacmmse ~ mean * adlike + age,
                data = naccmmse,
                prior = mmse_priors,
                family = "skew_normal",
                iter = 5000,
                chains = 4,
                seed = 5)

fit_mmse_summary <- summary(fit_mmse)
fit_mmse_summary
```


```{r}
# Plot posteriors
fit_mmse %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = fit_mmse_summary$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = fit_mmse_summary$fixed[2,4],
             colour = "blue",
             linetype = "dashed")
```


## piecewise linear regression model

```{r}
model_bp <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower = 0.25, upper = .75> bp ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < bp) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - bp) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - bp) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 bp ~ normal(0.5, 1) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 vector[25] sim_conditional_mean ;
 
 for (i in 1:25) {
  if ((i * 0.02) + 0.25 < bp) {
   sim_conditional_mean[i] = intercept + slope_before * (i - bp) ;
  } else {
   sim_conditional_mean[i] = intercept + slope_after * (i - bp) ;
  }
 }
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(naccmmse),
  clinscore = naccmmse$nacmmse,
  adlike = naccmmse$mean
)

fit_bp <- stan(model_code = model_bp,
               data = data_list)
```

```{r}
outmat <- summary(fit_bp,
                  par = c("intercept", "bp", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```



```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]
bp = outmat[["summary"]]["bp", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after
d = bp

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = d, yend = b) +
  geom_segment(x = d, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.4,
            label = paste("Breakpoint:", round(bp, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

# Fixed breakpoint

```{r}
model_bp_fixed <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < 0.5) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - 0.5) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - 0.5) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(naccmmse),
  clinscore = naccmmse$nacmmse,
  adlike = naccmmse$mean
)

fit_bp_fixed <- stan(model_code = model_bp_fixed,
                     data = data_list)
```

```{r}
outmat <- summary(fit_bp_fixed,
                  par = c("intercept", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```

```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = .5, yend = b) +
  geom_segment(x = .5, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```


```{r}
model_bp_linear <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 for (i in 1:N) {
  conditional_adlike[i] = intercept + slope * (adlike[i] - 0.5) ;
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(naccmmse),
  clinscore = naccmmse$nacmmse,
  adlike = naccmmse$mean
)

fit_bp_linear <- stan(model_code = model_bp_linear,
                      data = data_list)
```

```{r}
print(fit_bp_linear,
      par = c("intercept", "slope", "error"))
```

```{r}
# model comparison between fixed and unfixed breakpoint analysis
log_lik_fixed = extract_log_lik(fit_bp_fixed, merge_chains = FALSE)
r_eff_fixed = relative_eff(log_lik_fixed)
loo_fixed <- loo(log_lik_fixed, r_eff = r_eff_fixed)

log_lik_variable = extract_log_lik(fit_bp, merge_chains = FALSE)
r_eff_variable = relative_eff(log_lik_variable)
loo_variable <- loo(log_lik_variable, r_eff = r_eff_variable)

log_lik_linear = extract_log_lik(fit_bp_linear, merge_chains = FALSE)
r_eff_linear = relative_eff(log_lik_linear)
loo_linear <- loo(log_lik_linear, r_eff = r_eff_linear)

print(loo_fixed)
print(loo_variable)
print(loo_linear)

loo_compare(loo_fixed,
            loo_variable,
            loo_linear)
```

```{r}
rm(naccmmse)
```


# MOCA

```{r}
clinscore_name = "MoCA"

naccmoca_scores <- results %>%
  filter(!is.na(naccmoca))

naccmoca_summary <- naccmoca_scores %>%
  group_by(adlike) %>%
  summarise(n = length(subj),
            moca = paste(round(mean(naccmoca), digits = 1),
                         " (", round(sd(naccmoca), digits = 1), ")",
                         sep = ""),
            age = round(mean(age, na.rm = TRUE), digits = 1),
            sex = paste(length(gender[gender == "Female"]),
                        length(gender[gender == "Male"]),
                        sep = "/"))

datatable(naccmoca_summary)
```
```{r}
p <- ggplot(naccmoca_scores,
            aes(y = naccmoca,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
p_clinscore <- ggplot(naccmoca_scores, aes(x = mean, y = naccmoca, group = adlike, colour = adlike))
p_clinscore <- p_clinscore + geom_point() + geom_smooth(method = lm)
print(p_clinscore)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(naccmoca ~ adlike + age, 
                data = naccmoca_scores, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)


```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(naccmoca_scores[naccmoca_scores$adlike == "Negative", "naccmoca"][["naccmoca"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

```{r}
# z-score values
mean_age = mean(naccmoca_scores$age)
sd_age = sd(naccmoca_scores$age)
naccmoca_scores$age <- (naccmoca_scores$age - mean_age) / sd_age

mean_naccmoca = mean(naccmoca_scores$naccmoca)
sd_naccmoca = sd(naccmoca_scores$naccmoca)
naccmoca_scores$naccmoca <- (naccmoca_scores$naccmoca - mean_naccmoca) / sd_naccmoca

mean_mean = mean(naccmoca_scores$mean)
sd_mean = sd(naccmoca_scores$mean)
naccmoca_scores$mean <- (naccmoca_scores$mean - mean_mean) / sd_mean


# set priors
mmse_priors = c(set_prior("cauchy(0,1)", class = "b", coef = "age"),
                set_prior("cauchy(0,1)", class = "b", coef = "adlikePositive"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean:adlikePositive"))

fit_mmse <- brm(naccmoca ~ mean * adlike + age,
                data = naccmoca_scores,
                prior = mmse_priors,
                family = "skew_normal",
                iter = 5000,
                chains = 4,
                seed = 5)

fit_mmse_summary <- summary(fit_mmse)
fit_mmse_summary
```


```{r}
# Plot posteriors
fit_mmse %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = fit_mmse_summary$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = fit_mmse_summary$fixed[2,4],
             colour = "blue",
             linetype = "dashed")
```

## piecewise linear regression model

```{r}
model_bp <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower = 0.25, upper = .75> bp ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < bp) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - bp) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - bp) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 bp ~ normal(0.5, 1) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 vector[25] sim_conditional_mean ;
 
 for (i in 1:25) {
  if ((i * 0.02) + 0.25 < bp) {
   sim_conditional_mean[i] = intercept + slope_before * (i - bp) ;
  } else {
   sim_conditional_mean[i] = intercept + slope_after * (i - bp) ;
  }
 }
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(naccmoca_scores),
  clinscore = naccmoca_scores$naccmoca,
  adlike = naccmoca_scores$mean
)

fit_bp <- stan(model_code = model_bp,
               data = data_list)
```

```{r}
outmat <- summary(fit_bp,
                  par = c("intercept", "bp", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```



```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]
bp = outmat[["summary"]]["bp", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after
d = bp

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = d, yend = b) +
  geom_segment(x = d, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.4,
            label = paste("Breakpoint:", round(bp, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## Fixed breakpoint

```{r}
model_bp_fixed <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < 0.5) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - 0.5) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - 0.5) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(naccmoca_scores),
  clinscore = naccmoca_scores$naccmoca,
  adlike = naccmoca_scores$mean
)

fit_bp_fixed <- stan(model_code = model_bp_fixed,
                     data = data_list)
```

```{r}
outmat <- summary(fit_bp_fixed,
                  par = c("intercept", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```

```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = .5, yend = b) +
  geom_segment(x = .5, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## no breakpoint
```{r}
model_bp_linear <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 for (i in 1:N) {
  conditional_adlike[i] = intercept + slope * (adlike[i] - 0.5) ;
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(naccmoca_scores),
  clinscore = naccmoca_scores$naccmoca,
  adlike = naccmoca_scores$mean
)

fit_bp_linear <- stan(model_code = model_bp_linear,
                      data = data_list)
```

```{r}
print(fit_bp_linear,
      par = c("intercept", "slope", "error"))
```

```{r}
# model comparison between fixed and unfixed breakpoint analysis
log_lik_fixed = extract_log_lik(fit_bp_fixed, merge_chains = FALSE)
r_eff_fixed = relative_eff(log_lik_fixed)
loo_fixed <- loo(log_lik_fixed, r_eff = r_eff_fixed)

log_lik_variable = extract_log_lik(fit_bp, merge_chains = FALSE)
r_eff_variable = relative_eff(log_lik_variable)
loo_variable <- loo(log_lik_variable, r_eff = r_eff_variable)

log_lik_linear = extract_log_lik(fit_bp_linear, merge_chains = FALSE)
r_eff_linear = relative_eff(log_lik_linear)
loo_linear <- loo(log_lik_linear, r_eff = r_eff_linear)

print(loo_fixed)
print(loo_variable)
print(loo_linear)

loo_compare(loo_fixed,
            loo_variable,
            loo_linear)
```

```{r}
rm(naccmoca_scores)
```

# Digit span backwards 

```{r}
clinscore_name = "Backward digit span"

digib_scores <- results %>%
  filter(!is.na(digib))

digib_summary <- digib_scores %>%
  group_by(adlike) %>%
  summarise(n = length(subj),
            moca = paste(round(mean(digib), digits = 1),
                         " (", round(sd(digib), digits = 1), ")",
                         sep = ""),
            age = round(mean(age, na.rm = TRUE), digits = 1),
            sex = paste(length(gender[gender == "Female"]),
                        length(gender[gender == "Male"]),
                        sep = "/"))

datatable(digib_summary)
```
```{r}
p <- ggplot(digib_scores,
            aes(y = digib,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
p_clinscore <- ggplot(digib_scores, aes(x = mean, y = digib, group = adlike, colour = adlike))
p_clinscore <- p_clinscore + geom_point() + geom_smooth(method = lm)
print(p_clinscore)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(digib ~ adlike + age, 
                data = digib_scores, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)


```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(digib_scores[digib_scores$adlike == "Negative", "digib"][["digib"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

```{r}
# z-score values
mean_age = mean(digib_scores$age)
sd_age = sd(digib_scores$age)
digib_scores$age <- (digib_scores$age - mean_age) / sd_age

mean_digib = mean(digib_scores$digib)
sd_digib = sd(digib_scores$digib)
digib_scores$digib <- (digib_scores$digib - mean_digib) / sd_digib

mean_mean = mean(digib_scores$mean)
sd_mean = sd(digib_scores$mean)
digib_scores$mean <- (digib_scores$mean - mean_mean) / sd_mean


# set priors
mmse_priors = c(set_prior("cauchy(0,1)", class = "b", coef = "age"),
                set_prior("cauchy(0,1)", class = "b", coef = "adlikePositive"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean:adlikePositive"))

fit_mmse <- brm(digib ~ mean * adlike + age,
                data = digib_scores,
                prior = mmse_priors,
                family = "skew_normal",
                iter = 5000,
                chains = 4,
                seed = 5)

fit_mmse_summary <- summary(fit_mmse)
fit_mmse_summary
```


```{r}
# Plot posteriors
fit_mmse %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = fit_mmse_summary$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = fit_mmse_summary$fixed[2,4],
             colour = "blue",
             linetype = "dashed")
```

## piecewise linear regression model

```{r}
model_bp <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower = 0.25, upper = .75> bp ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < bp) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - bp) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - bp) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 bp ~ normal(0.5, 1) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 vector[25] sim_conditional_mean ;
 
 for (i in 1:25) {
  if ((i * 0.02) + 0.25 < bp) {
   sim_conditional_mean[i] = intercept + slope_before * (i - bp) ;
  } else {
   sim_conditional_mean[i] = intercept + slope_after * (i - bp) ;
  }
 }
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(digib_scores),
  clinscore = digib_scores$digib,
  adlike = digib_scores$mean
)

fit_bp <- stan(model_code = model_bp,
               data = data_list)
```

```{r}
outmat <- summary(fit_bp,
                  par = c("intercept", "bp", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```



```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]
bp = outmat[["summary"]]["bp", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after
d = bp

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = d, yend = b) +
  geom_segment(x = d, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.4,
            label = paste("Breakpoint:", round(bp, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## Fixed breakpoint

```{r}
model_bp_fixed <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < 0.5) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - 0.5) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - 0.5) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(digib_scores),
  clinscore = digib_scores$digib,
  adlike = digib_scores$mean
)

fit_bp_fixed <- stan(model_code = model_bp_fixed,
                     data = data_list)
```

```{r}
outmat <- summary(fit_bp_fixed,
                  par = c("intercept", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```

```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = .5, yend = b) +
  geom_segment(x = .5, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## no breakpoint
```{r}
model_bp_linear <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 for (i in 1:N) {
  conditional_adlike[i] = intercept + slope * (adlike[i] - 0.5) ;
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(digib_scores),
  clinscore = digib_scores$digib,
  adlike = digib_scores$mean
)

fit_bp_linear <- stan(model_code = model_bp_linear,
                      data = data_list)
```

```{r}
print(fit_bp_linear,
      par = c("intercept", "slope", "error"))
```

```{r}
# model comparison between fixed and unfixed breakpoint analysis
log_lik_fixed = extract_log_lik(fit_bp_fixed, merge_chains = FALSE)
r_eff_fixed = relative_eff(log_lik_fixed)
loo_fixed <- loo(log_lik_fixed, r_eff = r_eff_fixed)

log_lik_variable = extract_log_lik(fit_bp, merge_chains = FALSE)
r_eff_variable = relative_eff(log_lik_variable)
loo_variable <- loo(log_lik_variable, r_eff = r_eff_variable)

log_lik_linear = extract_log_lik(fit_bp_linear, merge_chains = FALSE)
r_eff_linear = relative_eff(log_lik_linear)
loo_linear <- loo(log_lik_linear, r_eff = r_eff_linear)

print(loo_fixed)
print(loo_variable)
print(loo_linear)

loo_compare(loo_fixed,
            loo_variable,
            loo_linear)
```

# Digit span forwards 

```{r}
clinscore_name = "Forward digit span"

digif_scores <- results %>%
  filter(!is.na(digif))

digif_summary <- digif_scores %>%
  group_by(adlike) %>%
  summarise(n = length(subj),
            moca = paste(round(mean(digif), digits = 1),
                         " (", round(sd(digif), digits = 1), ")",
                         sep = ""),
            age = round(mean(age, na.rm = TRUE), digits = 1),
            sex = paste(length(gender[gender == "Female"]),
                        length(gender[gender == "Male"]),
                        sep = "/"))

datatable(digif_summary)
```
```{r}
p <- ggplot(digif_scores,
            aes(y = digif,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
p_clinscore <- ggplot(digif_scores, aes(x = mean, y = digif, group = adlike, colour = adlike))
p_clinscore <- p_clinscore + geom_point() + geom_smooth(method = lm)
print(p_clinscore)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(digif ~ adlike + age, 
                data = digif_scores, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)


```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(digif_scores[digif_scores$adlike == "Negative", "digif"][["digif"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

```{r}
# z-score values
mean_age = mean(digif_scores$age)
sd_age = sd(digif_scores$age)
digif_scores$age <- (digif_scores$age - mean_age) / sd_age

mean_digif = mean(digif_scores$digif)
sd_digif = sd(digif_scores$digif)
digif_scores$digif <- (digif_scores$digif - mean_digif) / sd_digif

mean_mean = mean(digif_scores$mean)
sd_mean = sd(digif_scores$mean)
digif_scores$mean <- (digif_scores$mean - mean_mean) / sd_mean


# set priors
mmse_priors = c(set_prior("cauchy(0,1)", class = "b", coef = "age"),
                set_prior("cauchy(0,1)", class = "b", coef = "adlikePositive"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean:adlikePositive"))

fit_mmse <- brm(digif ~ mean * adlike + age,
                data = digif_scores,
                prior = mmse_priors,
                family = "skew_normal",
                iter = 5000,
                chains = 4,
                seed = 5)

fit_mmse_summary <- summary(fit_mmse)
fit_mmse_summary
```


```{r}
# Plot posteriors
fit_mmse %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = fit_mmse_summary$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = fit_mmse_summary$fixed[2,4],
             colour = "blue",
             linetype = "dashed")
```

## piecewise linear regression model

```{r}
model_bp <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower = 0.25, upper = .75> bp ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < bp) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - bp) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - bp) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 bp ~ normal(0.5, 1) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 vector[25] sim_conditional_mean ;
 
 for (i in 1:25) {
  if ((i * 0.02) + 0.25 < bp) {
   sim_conditional_mean[i] = intercept + slope_before * (i - bp) ;
  } else {
   sim_conditional_mean[i] = intercept + slope_after * (i - bp) ;
  }
 }
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(digif_scores),
  clinscore = digif_scores$digif,
  adlike = digif_scores$mean
)

fit_bp <- stan(model_code = model_bp,
               data = data_list)
```

```{r}
outmat <- summary(fit_bp,
                  par = c("intercept", "bp", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```



```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]
bp = outmat[["summary"]]["bp", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after
d = bp

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = d, yend = b) +
  geom_segment(x = d, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.4,
            label = paste("Breakpoint:", round(bp, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## Fixed breakpoint

```{r}
model_bp_fixed <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < 0.5) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - 0.5) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - 0.5) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(digif_scores),
  clinscore = digif_scores$digif,
  adlike = digif_scores$mean
)

fit_bp_fixed <- stan(model_code = model_bp_fixed,
                     data = data_list)
```

```{r}
outmat <- summary(fit_bp_fixed,
                  par = c("intercept", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```

```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = .5, yend = b) +
  geom_segment(x = .5, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## no breakpoint
```{r}
model_bp_linear <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 for (i in 1:N) {
  conditional_adlike[i] = intercept + slope * (adlike[i] - 0.5) ;
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(digif_scores),
  clinscore = digif_scores$digif,
  adlike = digif_scores$mean
)

fit_bp_linear <- stan(model_code = model_bp_linear,
                      data = data_list)
```

```{r}
print(fit_bp_linear,
      par = c("intercept", "slope", "error"))
```

```{r}
# model comparison between fixed and unfixed breakpoint analysis
log_lik_fixed = extract_log_lik(fit_bp_fixed, merge_chains = FALSE)
r_eff_fixed = relative_eff(log_lik_fixed)
loo_fixed <- loo(log_lik_fixed, r_eff = r_eff_fixed)

log_lik_variable = extract_log_lik(fit_bp, merge_chains = FALSE)
r_eff_variable = relative_eff(log_lik_variable)
loo_variable <- loo(log_lik_variable, r_eff = r_eff_variable)

log_lik_linear = extract_log_lik(fit_bp_linear, merge_chains = FALSE)
r_eff_linear = relative_eff(log_lik_linear)
loo_linear <- loo(log_lik_linear, r_eff = r_eff_linear)

print(loo_fixed)
print(loo_variable)
print(loo_linear)

loo_compare(loo_fixed,
            loo_variable,
            loo_linear)
```

# Animal fluency

```{r}
clinscore_name = "Category fluency"

animals_scores <- results %>%
  filter(!is.na(animals))

animals_summary <- animals_scores %>%
  group_by(adlike) %>%
  summarise(n = length(subj),
            moca = paste(round(mean(animals), digits = 1),
                         " (", round(sd(animals), digits = 1), ")",
                         sep = ""),
            age = round(mean(age, na.rm = TRUE), digits = 1),
            sex = paste(length(gender[gender == "Female"]),
                        length(gender[gender == "Male"]),
                        sep = "/"))

datatable(animals_summary)
```
```{r}
p <- ggplot(animals_scores,
            aes(y = animals,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
p_clinscore <- ggplot(animals_scores, aes(x = mean, y = animals, group = adlike, colour = adlike))
p_clinscore <- p_clinscore + geom_point() + geom_smooth(method = lm)
print(p_clinscore)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(animals ~ adlike + age, 
                data = animals_scores, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)


```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(animals_scores[animals_scores$adlike == "Negative", "animals"][["animals"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

```{r}
# z-score values
mean_age = mean(animals_scores$age)
sd_age = sd(animals_scores$age)
animals_scores$age <- (animals_scores$age - mean_age) / sd_age

mean_animals = mean(animals_scores$animals)
sd_animals = sd(animals_scores$animals)
animals_scores$animals <- (animals_scores$animals - mean_animals) / sd_animals

mean_mean = mean(animals_scores$mean)
sd_mean = sd(animals_scores$mean)
animals_scores$mean <- (animals_scores$mean - mean_mean) / sd_mean


# set priors
mmse_priors = c(set_prior("cauchy(0,1)", class = "b", coef = "age"),
                set_prior("cauchy(0,1)", class = "b", coef = "adlikePositive"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean:adlikePositive"))

fit_mmse <- brm(animals ~ mean * adlike + age,
                data = animals_scores,
                prior = mmse_priors,
                family = "skew_normal",
                iter = 5000,
                chains = 4,
                seed = 5)

fit_mmse_summary <- summary(fit_mmse)
fit_mmse_summary
```


```{r}
# Plot posteriors
fit_mmse %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = fit_mmse_summary$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = fit_mmse_summary$fixed[2,4],
             colour = "blue",
             linetype = "dashed")
```

## piecewise linear regression model

```{r}
model_bp <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower = 0.25, upper = .75> bp ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < bp) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - bp) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - bp) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 bp ~ normal(0.5, 1) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 vector[25] sim_conditional_mean ;
 
 for (i in 1:25) {
  if ((i * 0.02) + 0.25 < bp) {
   sim_conditional_mean[i] = intercept + slope_before * (i - bp) ;
  } else {
   sim_conditional_mean[i] = intercept + slope_after * (i - bp) ;
  }
 }
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(animals_scores),
  clinscore = animals_scores$animals,
  adlike = animals_scores$mean
)

fit_bp <- stan(model_code = model_bp,
               data = data_list)
```

```{r}
outmat <- summary(fit_bp,
                  par = c("intercept", "bp", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```



```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]
bp = outmat[["summary"]]["bp", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after
d = bp

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = d, yend = b) +
  geom_segment(x = d, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.4,
            label = paste("Breakpoint:", round(bp, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## Fixed breakpoint

```{r}
model_bp_fixed <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < 0.5) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - 0.5) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - 0.5) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(animals_scores),
  clinscore = animals_scores$animals,
  adlike = animals_scores$mean
)

fit_bp_fixed <- stan(model_code = model_bp_fixed,
                     data = data_list)
```

```{r}
outmat <- summary(fit_bp_fixed,
                  par = c("intercept", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```

```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = .5, yend = b) +
  geom_segment(x = .5, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## no breakpoint
```{r}
model_bp_linear <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 for (i in 1:N) {
  conditional_adlike[i] = intercept + slope * (adlike[i] - 0.5) ;
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(animals_scores),
  clinscore = animals_scores$animals,
  adlike = animals_scores$mean
)

fit_bp_linear <- stan(model_code = model_bp_linear,
                      data = data_list)
```

```{r}
print(fit_bp_linear,
      par = c("intercept", "slope", "error"))
```

```{r}
# model comparison between fixed and unfixed breakpoint analysis
log_lik_fixed = extract_log_lik(fit_bp_fixed, merge_chains = FALSE)
r_eff_fixed = relative_eff(log_lik_fixed)
loo_fixed <- loo(log_lik_fixed, r_eff = r_eff_fixed)

log_lik_variable = extract_log_lik(fit_bp, merge_chains = FALSE)
r_eff_variable = relative_eff(log_lik_variable)
loo_variable <- loo(log_lik_variable, r_eff = r_eff_variable)

log_lik_linear = extract_log_lik(fit_bp_linear, merge_chains = FALSE)
r_eff_linear = relative_eff(log_lik_linear)
loo_linear <- loo(log_lik_linear, r_eff = r_eff_linear)

print(loo_fixed)
print(loo_variable)
print(loo_linear)

loo_compare(loo_fixed,
            loo_variable,
            loo_linear)
```

# Trails B

```{r}
clinscore_name = "Trails B"
trialsb_scores <- results %>%
  filter(!is.na(trialsb))

trialsb_summary <- trialsb_scores %>%
  group_by(adlike) %>%
  summarise(n = length(subj),
            moca = paste(round(mean(trialsb), digits = 1),
                         " (", round(sd(trialsb), digits = 1), ")",
                         sep = ""),
            age = round(mean(age, na.rm = TRUE), digits = 1),
            sex = paste(length(gender[gender == "Female"]),
                        length(gender[gender == "Male"]),
                        sep = "/"))

datatable(trialsb_summary)
```
```{r}
p <- ggplot(trialsb_scores,
            aes(y = trialsb,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
p_clinscore <- ggplot(trialsb_scores, aes(x = mean, y = trialsb, group = adlike, colour = adlike))
p_clinscore <- p_clinscore + geom_point() + geom_smooth(method = lm)
print(p_clinscore)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(trialsb ~ adlike + age, 
                data = trialsb_scores, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)


```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(trialsb_scores[trialsb_scores$adlike == "Negative", "trialsb"][["trialsb"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

```{r}
# z-score values
mean_age = mean(trialsb_scores$age)
sd_age = sd(trialsb_scores$age)
trialsb_scores$age <- (trialsb_scores$age - mean_age) / sd_age

mean_trialsb = mean(trialsb_scores$trialsb)
sd_trialsb = sd(trialsb_scores$trialsb)
trialsb_scores$trialsb <- (trialsb_scores$trialsb - mean_trialsb) / sd_trialsb

mean_mean = mean(trialsb_scores$mean)
sd_mean = sd(trialsb_scores$mean)
trialsb_scores$mean <- (trialsb_scores$mean - mean_mean) / sd_mean


# set priors
mmse_priors = c(set_prior("cauchy(0,1)", class = "b", coef = "age"),
                set_prior("cauchy(0,1)", class = "b", coef = "adlikePositive"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean:adlikePositive"))

fit_mmse <- brm(trialsb ~ mean * adlike + age,
                data = trialsb_scores,
                prior = mmse_priors,
                family = "skew_normal",
                iter = 5000,
                chains = 4,
                seed = 5)

fit_mmse_summary <- summary(fit_mmse)
fit_mmse_summary
```


```{r}
# Plot posteriors
fit_mmse %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = fit_mmse_summary$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = fit_mmse_summary$fixed[2,4],
             colour = "blue",
             linetype = "dashed")
```

## piecewise linear regression model

```{r}
model_bp <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower = 0.25, upper = .75> bp ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < bp) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - bp) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - bp) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 bp ~ normal(0.5, 1) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 vector[25] sim_conditional_mean ;
 
 for (i in 1:25) {
  if ((i * 0.02) + 0.25 < bp) {
   sim_conditional_mean[i] = intercept + slope_before * (i - bp) ;
  } else {
   sim_conditional_mean[i] = intercept + slope_after * (i - bp) ;
  }
 }
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(trialsb_scores),
  clinscore = trialsb_scores$trialsb,
  adlike = trialsb_scores$mean
)

fit_bp <- stan(model_code = model_bp,
               data = data_list)
```

```{r}
outmat <- summary(fit_bp,
                  par = c("intercept", "bp", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```



```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]
bp = outmat[["summary"]]["bp", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after
d = bp

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = d, yend = b) +
  geom_segment(x = d, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.4,
            label = paste("Breakpoint:", round(bp, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## Fixed breakpoint

```{r}
model_bp_fixed <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < 0.5) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - 0.5) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - 0.5) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(trialsb_scores),
  clinscore = trialsb_scores$trialsb,
  adlike = trialsb_scores$mean
)

fit_bp_fixed <- stan(model_code = model_bp_fixed,
                     data = data_list)
```

```{r}
outmat <- summary(fit_bp_fixed,
                  par = c("intercept", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```

```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = .5, yend = b) +
  geom_segment(x = .5, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## no breakpoint
```{r}
model_bp_linear <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 for (i in 1:N) {
  conditional_adlike[i] = intercept + slope * (adlike[i] - 0.5) ;
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(trialsb_scores),
  clinscore = trialsb_scores$trialsb,
  adlike = trialsb_scores$mean
)

fit_bp_linear <- stan(model_code = model_bp_linear,
                      data = data_list)
```

```{r}
print(fit_bp_linear,
      par = c("intercept", "slope", "error"))
```

```{r}
# model comparison between fixed and unfixed breakpoint analysis
log_lik_fixed = extract_log_lik(fit_bp_fixed, merge_chains = FALSE)
r_eff_fixed = relative_eff(log_lik_fixed)
loo_fixed <- loo(log_lik_fixed, r_eff = r_eff_fixed)

log_lik_variable = extract_log_lik(fit_bp, merge_chains = FALSE)
r_eff_variable = relative_eff(log_lik_variable)
loo_variable <- loo(log_lik_variable, r_eff = r_eff_variable)

log_lik_linear = extract_log_lik(fit_bp_linear, merge_chains = FALSE)
r_eff_linear = relative_eff(log_lik_linear)
loo_linear <- loo(log_lik_linear, r_eff = r_eff_linear)

print(loo_fixed)
print(loo_variable)
print(loo_linear)

loo_compare(loo_fixed,
            loo_variable,
            loo_linear)
```


# WAIS

```{r}
clinscore_name = "WAIS"

wais_scores <- results %>%
  filter(!is.na(wais))

wais_summary <- wais_scores %>%
  group_by(adlike) %>%
  summarise(n = length(subj),
            moca = paste(round(mean(wais), digits = 1),
                         " (", round(sd(wais), digits = 1), ")",
                         sep = ""),
            age = round(mean(age, na.rm = TRUE), digits = 1),
            sex = paste(length(gender[gender == "Female"]),
                        length(gender[gender == "Male"]),
                        sep = "/"))

datatable(wais_summary)
```
```{r}
p <- ggplot(wais_scores,
            aes(y = wais,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
p_clinscore <- ggplot(wais_scores, aes(x = mean, y = wais, group = adlike, colour = adlike))
p_clinscore <- p_clinscore + geom_point() + geom_smooth(method = lm)
print(p_clinscore)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(wais ~ adlike + age, 
                data = wais_scores, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)


```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(wais_scores[wais_scores$adlike == "Negative", "wais"][["wais"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

```{r}
# z-score values
mean_age = mean(wais_scores$age)
sd_age = sd(wais_scores$age)
wais_scores$age <- (wais_scores$age - mean_age) / sd_age

mean_wais = mean(wais_scores$wais)
sd_wais = sd(wais_scores$wais)
wais_scores$wais <- (wais_scores$wais - mean_wais) / sd_wais

mean_mean = mean(wais_scores$mean)
sd_mean = sd(wais_scores$mean)
wais_scores$mean <- (wais_scores$mean - mean_mean) / sd_mean


# set priors
mmse_priors = c(set_prior("cauchy(0,1)", class = "b", coef = "age"),
                set_prior("cauchy(0,1)", class = "b", coef = "adlikePositive"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean:adlikePositive"))

fit_mmse <- brm(wais ~ mean * adlike + age,
                data = wais_scores,
                prior = mmse_priors,
                family = "skew_normal",
                iter = 5000,
                chains = 4,
                seed = 5)

fit_mmse_summary <- summary(fit_mmse)
fit_mmse_summary
```


```{r}
# Plot posteriors
fit_mmse %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = fit_mmse_summary$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = fit_mmse_summary$fixed[2,4],
             colour = "blue",
             linetype = "dashed")
```

## piecewise linear regression model

```{r}
model_bp <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower = 0.25, upper = .75> bp ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < bp) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - bp) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - bp) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 bp ~ normal(0.5, 1) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 vector[25] sim_conditional_mean ;
 
 for (i in 1:25) {
  if ((i * 0.02) + 0.25 < bp) {
   sim_conditional_mean[i] = intercept + slope_before * (i - bp) ;
  } else {
   sim_conditional_mean[i] = intercept + slope_after * (i - bp) ;
  }
 }
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(wais_scores),
  clinscore = wais_scores$wais,
  adlike = wais_scores$mean
)

fit_bp <- stan(model_code = model_bp,
               data = data_list)
```

```{r}
outmat <- summary(fit_bp,
                  par = c("intercept", "bp", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```



```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]
bp = outmat[["summary"]]["bp", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after
d = bp

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = d, yend = b) +
  geom_segment(x = d, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.4,
            label = paste("Breakpoint:", round(bp, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## Fixed breakpoint

```{r}
model_bp_fixed <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < 0.5) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - 0.5) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - 0.5) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(wais_scores),
  clinscore = wais_scores$wais,
  adlike = wais_scores$mean
)

fit_bp_fixed <- stan(model_code = model_bp_fixed,
                     data = data_list)
```

```{r}
outmat <- summary(fit_bp_fixed,
                  par = c("intercept", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```

```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = .5, yend = b) +
  geom_segment(x = .5, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## no breakpoint
```{r}
model_bp_linear <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 for (i in 1:N) {
  conditional_adlike[i] = intercept + slope * (adlike[i] - 0.5) ;
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(wais_scores),
  clinscore = wais_scores$wais,
  adlike = wais_scores$mean
)

fit_bp_linear <- stan(model_code = model_bp_linear,
                      data = data_list)
```

```{r}
print(fit_bp_linear,
      par = c("intercept", "slope", "error"))
```

```{r}
# model comparison between fixed and unfixed breakpoint analysis
log_lik_fixed = extract_log_lik(fit_bp_fixed, merge_chains = FALSE)
r_eff_fixed = relative_eff(log_lik_fixed)
loo_fixed <- loo(log_lik_fixed, r_eff = r_eff_fixed)

log_lik_variable = extract_log_lik(fit_bp, merge_chains = FALSE)
r_eff_variable = relative_eff(log_lik_variable)
loo_variable <- loo(log_lik_variable, r_eff = r_eff_variable)

log_lik_linear = extract_log_lik(fit_bp_linear, merge_chains = FALSE)
r_eff_linear = relative_eff(log_lik_linear)
loo_linear <- loo(log_lik_linear, r_eff = r_eff_linear)

print(loo_fixed)
print(loo_variable)
print(loo_linear)

loo_compare(loo_fixed,
            loo_variable,
            loo_linear)
```


# Boston naming task

```{r}
clinscore_name <- "Boston naming task"
boston_scores <- results %>%
  filter(!is.na(boston))

boston_summary <- boston_scores %>%
  group_by(adlike) %>%
  summarise(n = length(subj),
            moca = paste(round(mean(boston), digits = 1),
                         " (", round(sd(boston), digits = 1), ")",
                         sep = ""),
            age = round(mean(age, na.rm = TRUE), digits = 1),
            sex = paste(length(gender[gender == "Female"]),
                        length(gender[gender == "Male"]),
                        sep = "/"))

datatable(boston_summary)
```
```{r}
p <- ggplot(boston_scores,
            aes(y = boston,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
p_clinscore <- ggplot(boston_scores, aes(x = mean, y = boston, group = adlike, colour = adlike))
p_clinscore <- p_clinscore + geom_point() + geom_smooth(method = lm)
print(p_clinscore)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(boston ~ adlike + age, 
                data = boston_scores, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)


```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(boston_scores[boston_scores$adlike == "Negative", "boston"][["boston"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

```{r}
# z-score values
mean_age = mean(boston_scores$age)
sd_age = sd(boston_scores$age)
boston_scores$age <- (boston_scores$age - mean_age) / sd_age

mean_boston = mean(boston_scores$boston)
sd_boston = sd(boston_scores$boston)
boston_scores$boston <- (boston_scores$boston - mean_boston) / sd_boston

mean_mean = mean(boston_scores$mean)
sd_mean = sd(boston_scores$mean)
boston_scores$mean <- (boston_scores$mean - mean_mean) / sd_mean


# set priors
mmse_priors = c(set_prior("cauchy(0,1)", class = "b", coef = "age"),
                set_prior("cauchy(0,1)", class = "b", coef = "adlikePositive"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean"),
                set_prior("cauchy(0,1)", class = "b", coef = "mean:adlikePositive"))

fit_mmse <- brm(boston ~ mean * adlike + age,
                data = boston_scores,
                prior = mmse_priors,
                family = "skew_normal",
                iter = 5000,
                chains = 4,
                seed = 5)

fit_mmse_summary <- summary(fit_mmse)
fit_mmse_summary
```


```{r}
# Plot posteriors
fit_mmse %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = fit_mmse_summary$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = fit_mmse_summary$fixed[2,4],
             colour = "blue",
             linetype = "dashed")
```

## piecewise linear regression model

```{r}
model_bp <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower = 0.25, upper = .75> bp ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < bp) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - bp) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - bp) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 bp ~ normal(0.5, 1) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 vector[25] sim_conditional_mean ;
 
 for (i in 1:25) {
  if ((i * 0.02) + 0.25 < bp) {
   sim_conditional_mean[i] = intercept + slope_before * (i - bp) ;
  } else {
   sim_conditional_mean[i] = intercept + slope_after * (i - bp) ;
  }
 }
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(boston_scores),
  clinscore = boston_scores$boston,
  adlike = boston_scores$mean
)

fit_bp <- stan(model_code = model_bp,
               data = data_list)
```

```{r}
outmat <- summary(fit_bp,
                  par = c("intercept", "bp", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```



```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]
bp = outmat[["summary"]]["bp", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after
d = bp

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = d, yend = b) +
  geom_segment(x = d, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.4,
            label = paste("Breakpoint:", round(bp, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## Fixed breakpoint

```{r}
model_bp_fixed <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope_before ;
 real slope_after ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 real slope_difference ;
 slope_difference = slope_after - slope_before ;
 for (i in 1:N) {
  if (adlike[i] < 0.5) {
   conditional_adlike[i] = intercept + slope_before * (adlike[i] - 0.5) ;
  } else {
   conditional_adlike[i] = intercept + slope_after * (adlike[i] - 0.5) ;
  }
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope_before ~ cauchy(0, 10) ;
 slope_after ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(boston_scores),
  clinscore = boston_scores$boston,
  adlike = boston_scores$mean
)

fit_bp_fixed <- stan(model_code = model_bp_fixed,
                     data = data_list)
```

```{r}
outmat <- summary(fit_bp_fixed,
                  par = c("intercept", "slope_before", "slope_after", "slope_difference", "error"))

outmat_summary <- xtable(outmat$summary,
                 display = c("s", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "fg", "f", "f"),
                 digits = c(0, 2, 2, 2, 2, 2, 2, 2, 2, 0, 2))

print(outmat_summary)
```

```{r}
slope_before = outmat[["summary"]]["slope_before", "mean"]
slope_after = outmat[["summary"]]["slope_after", "mean"]
intercept = outmat[["summary"]]["intercept", "mean"]

a = intercept
b = intercept + (0.5 * slope_before)
c = intercept + slope_after

df_blank = data.frame(x = c(-.5), y = c(0))
p <- ggplot(df_blank, aes(x = x, y = y)) + geom_point() + xlim(0, 1.) + ylim(bp_y_min, bp_y_max) +
  geom_segment(x = 0, y = a,
               xend = .5, yend = b) +
  geom_segment(x = .5, y = b,
               xend = 1., yend = c) +
  labs(x = "AD score",
       y = clinscore_name) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.25,
            label = paste("Slope before:", round(slope_before, digits = 2)),
            hjust = "left",
            size = 8) +
  geom_text(x = 0.,
            y = (bp_y_min) + ((bp_y_max) - (bp_y_min)) * 0.1,
            label = paste("Slope after:", round(slope_after, digits = 2)),
            hjust = "left",
            size = 8) +
  theme(axis.title = element_text(size = 20),
        axis.text = element_text(size = 16))

print(p)
```

## no breakpoint
```{r}
model_bp_linear <- '
data {
 int<lower=1> N ;
 real clinscore[N] ;
 real adlike[N] ;
}

parameters {
 real intercept ;
 real slope ;
 real<lower=0> error ;
}

transformed parameters {
 vector[N] conditional_adlike ;
 for (i in 1:N) {
  conditional_adlike[i] = intercept + slope * (adlike[i] - 0.5) ;
 }
}

model {
 //set priors
 intercept ~ cauchy(0, 10) ;
 slope ~ cauchy(0, 10) ;
 error ~ cauchy(0, 10) ;
 for (i in 1:N) {
  clinscore[i] ~ normal(conditional_adlike[i], error) ;
 }
}

generated quantities {
 vector[N] sim_clinscore ;
 vector[N] log_lik ;
 
 for (i in 1:N) {
  sim_clinscore[i] = normal_rng(conditional_adlike[i], error) ;
  log_lik[i] = normal_lpdf(clinscore[i] | conditional_adlike[i], error) ;
 }
}
'
```

```{r}
data_list <- list(
  N = nrow(boston_scores),
  clinscore = boston_scores$boston,
  adlike = boston_scores$mean
)

fit_bp_linear <- stan(model_code = model_bp_linear,
                      data = data_list)
```

```{r}
print(fit_bp_linear,
      par = c("intercept", "slope", "error"))
```

```{r}
# model comparison between fixed and unfixed breakpoint analysis
log_lik_fixed = extract_log_lik(fit_bp_fixed, merge_chains = FALSE)
r_eff_fixed = relative_eff(log_lik_fixed)
loo_fixed <- loo(log_lik_fixed, r_eff = r_eff_fixed)

log_lik_variable = extract_log_lik(fit_bp, merge_chains = FALSE)
r_eff_variable = relative_eff(log_lik_variable)
loo_variable <- loo(log_lik_variable, r_eff = r_eff_variable)

log_lik_linear = extract_log_lik(fit_bp_linear, merge_chains = FALSE)
r_eff_linear = relative_eff(log_lik_linear)
loo_linear <- loo(log_lik_linear, r_eff = r_eff_linear)

print(loo_fixed)
print(loo_variable)
print(loo_linear)

loo_compare(loo_fixed,
            loo_variable,
            loo_linear)
```




