---
title: "UKBB clinical scores"
output: html_notebook
---

```{r}
library(dplyr)
library(RSQLite)
library(ggplot2)
library(broom.mixed)
library(tidyr)
library(brms)
library(bayesplot)
library(rstan)
library(multcomp)
library(coda)
library(tidybayes)
library(gridExtra)

# options for rstan
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
set.seed(123)
```

In this notebook we'll examine the correlation between the neuroimaging derived phenotype scores and clinical scores in the UK biobank population.

```{r}
# create sqlite connection
ukbb.db <- DBI::dbConnect(SQLite(),
                          "../data/ukb40183.db")
```

```{r}
# define clinical score
threshold = 0.5

getdata <- function(clinscore, threshold){
  # pairs matching
  alldata <- dbGetQuery(ukbb.db,
                        sql(paste("SELECT clinicaldata.eid, clinicaldata.noncancer_code_selfreported,",
                                  "demogs.age, demogs.visit,",
                                  paste("clinicaldata", clinscore, sep = "."), " AS clinscore,",
                                  "ukbb_t1_phenotype.mean",
                                  "FROM clinicaldata",
                                  "JOIN ukbb_t1_phenotype, demogs, baseline ON clinicaldata.eid = ukbb_t1_phenotype.eid",
                                  "AND clinicaldata.eid = demogs.eid",
                                  "AND clinicaldata.visit = demogs.visit",
                                  "AND clinicaldata.eid = baseline.eid",
                                  "AND ukbb_t1_phenotype.eid = baseline.eid",
                                  "AND demogs.eid = baseline.eid",
                                  "AND ukbb_t1_phenotype.eid = demogs.eid",
                                  "WHERE demogs.age IS NOT ''",
                                  "AND baseline.imaging = 1",
                                  "AND", paste("clinicaldata", clinscore, sep = "."), "IS NOT ''"))) %>%
    filter(visit <= 2) %>%
    group_by(eid) %>%
    filter(visit == max(visit))

  alldata$adlike <- sapply(alldata$mean,
                           function(x, thr) if (x >=  thr) { return("Positive") } else {return("Negative")},
                           thr = threshold)
  
  alldata <- alldata %>%
    mutate(adlike = ifelse(noncancer_code_selfreported == "1263", "AD", adlike)) %>%
    filter(adlike != "AD") # filter out those with AD at recruitment
  
  alldata$adlike <- as.factor(alldata$adlike)
  alldata$clinscore <- as.numeric(alldata$clinscore)
  alldata$age <- as.numeric(alldata$age)
  
  # normalise age
  mean_age <- mean(alldata$age)
  sd_age <- sd(alldata$age)
  alldata$age <- (alldata$age - mean_age) / sd_age
  
  return(alldata)
}

```

# Cognitive tests

## Tower rearranging

```{r}
clinscore = "towerrearranging_n_correct"
clinscore_name = "Tower rearranging, n correct"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(10, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

## Pairs matching

### n correct
```{r}
clinscore = "pairsmatching_n_correct"
clinscore_name = "Pairs matching\nn correct"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold) %>%
  filter(clinscore <=3)

alldata$clinscore <- factor(alldata$clinscore + 1,
                            ordered = c(1,2,3,4))
  
alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid)) %>%
  filter(visit == 0)

```

```{r}
# change to proportions
total_pos = length(alldata[alldata$adlike == "Positive", "eid"][["eid"]])
total_neg = length(alldata[alldata$adlike == "Negative", "eid"][["eid"]])

alldata_summary <- alldata %>%
  group_by(adlike, clinscore) %>%
  summarise(n = length(eid))
alldata_summary$totals <- sapply(alldata_summary$adlike,
                                 function(x, total_pos, total_neg) if (x == "Positive") {return(total_pos)} else {return(total_neg)},
                                 total_pos = total_pos,
                                 total_neg = total_neg)
alldata_summary <- alldata_summary %>%
  group_by(adlike, clinscore) %>%
  mutate(proportion = n / totals)

p <- ggplot(alldata_summary,
            aes(y = proportion,
                x = adlike,
                fill = clinscore)) +
  geom_histogram(stat = "identity",
                 position = position_dodge2()) +
  labs(x = clinscore,
       fill = clinscore_name) +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```


```{r}
# priors
priors <- c(prior(student_t(3,0,2.5), class = "Intercept"))

# Bayesian analysis
cog_brms <- brm(clinscore ~  1 + cs(adlike) + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4,
                family = acat("probit"),
                prior = priors)

summary(cog_brms)
```

```{r}
ce <- conditional_effects(cog_brms, "adlike", categorical = TRUE)

p2 <- plot(ce, plot = FALSE)[[1]] +
  ylim(0.975, .995) +
  theme(legend.position = "None",
        axis.title.x = element_blank())

p3 <- plot(ce, plot = FALSE)[[1]] +
  ylim(0., .02) +
  theme(legend.position = "None",
        axis.title.x = element_blank())

p1 <- plot(ce, plot = FALSE)[[1]] +
  labs(x = "AD-like") +
  scale_fill_discrete(name = clinscore_name,
                      labels = seq(0, 3, by = 1)) +
  scale_colour_discrete(name = clinscore_name,
                        labels = seq(0, 3, by = 1))

grid.arrange(grobs = list(p1, p2, p3),
             widths = c(2,1),
             layout_matrix = rbind(c(1,2),
                                   c(1,3)))

```


```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[4,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[4,4],
             colour = "blue",
             linetype = "dashed")

```


### n incorrect
```{r}
clinscore = "pairsmatching_n_incorrect"
clinscore_name = "Pairs matching, n incorrect"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)
alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)

```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

### time to complete
```{r}
clinscore = "pairsmatching_time_to_complete"
clinscore_name = "Pairs matching, time to completion"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)
alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}

p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

## Reaction time

### Duration to snap button pressed
```{r}
clinscore = "reactiontime_duration_to_snapbutton_pressed"
clinscore_name = "Reaction time, duration to button press"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)
alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}

p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

### Mean time correct
```{r}
clinscore = "reactiontime_mean_time_correct"
clinscore_name = "Reaction time, mean time of correct trials"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)
alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}

p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

## Numeric memory
```{r}
clinscore = "nummemory_max_remembered"
clinscore_name = "Number memory"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)
alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}

p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

## Prospective memory

# result
```{r}
clinscore = "prospectivememory_result"
clinscore_name = "Prospective memory"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold) %>%
  filter(clinscore != 0)

alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# change to proportions
total_pos = length(alldata[alldata$adlike == "Positive", "eid"][["eid"]])
total_neg = length(alldata[alldata$adlike == "Negative", "eid"][["eid"]])

alldata_summary <- alldata %>%
  group_by(adlike, clinscore) %>%
  summarise(n = length(eid))
alldata_summary$totals <- sapply(alldata_summary$adlike,
                                 function(x, total_pos, total_neg) if (x == "Positive") {return(total_pos)} else {return(total_neg)},
                                 total_pos = total_pos,
                                 total_neg = total_neg)
alldata_summary <- alldata_summary %>%
  group_by(adlike, clinscore) %>%
  mutate(proportion = n / totals)

alldata_summary$clinscore <- as.integer(alldata_summary$clinscore)

p <- ggplot(alldata_summary,
            aes(y = proportion,
                x = adlike,
                fill = clinscore)) +
  geom_histogram(stat = "identity",
                 position = position_dodge2()) +
  labs(x = clinscore,
       fill = clinscore_name) +
  theme(axis.title.x = element_blank(),
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                family = binomial(link = "logit"),
                iter = 5000, warmup = 1000, chains = 4)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

## Fluid intelligence
```{r}
clinscore = "FI_total"
clinscore_name = "Fluid intelligence"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

## Matrix pattern completion
```{r}
clinscore = "mpc_n_solved"
clinscore_name = "Matrix pattern completion, n solved"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

## Trail making
### Path 2 errors
```{r}
clinscore = "trails_path_2_errors"
clinscore_name = "Trails B errors"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```

### Path 2 duration
```{r}
clinscore = "trails_path_2_duration"
clinscore_name = "Trails B duration"

alldata <- getdata(clinscore = clinscore,
                   threshold = threshold)

alldata %>%
  group_by(adlike, visit) %>%
  summarise(n = length(eid))
```

```{r}
p <- ggplot(alldata,
            aes(y = clinscore,
                x = adlike,
                fill = adlike)) +
  geom_boxplot() +
  labs(y = clinscore_name) +
  theme(axis.title.x = element_blank(),
        legend.position = "None",
        axis.text = element_text(size = 20),
        axis.title = element_text(size = 20))

print(p)
```

```{r}
# Bayesian analysis
priors <- c(prior(cauchy(0, 10), class = "Intercept"), 
            prior(cauchy(0, 10), class = "b"), 
            prior(cauchy(0, 10), class = "sigma"))

cog_brms <- brm(clinscore ~ adlike + age, 
                data = alldata, 
                iter = 5000, warmup = 1000, chains = 4, 
                prior = priors)

summary(cog_brms)

```

```{r}
summary_model <- summary(cog_brms)

cog_brms %>%
  spread_draws(b_adlikePositive) %>%
  ggplot(aes(x = b_adlikePositive)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(b_adlikePositive)),
             colour = "blue",
             linetype = "dashed") +
  geom_vline(xintercept = summary_model$fixed[2,3],
             colour = "blue",
             linetype = "dashed") + 
  geom_vline(xintercept = summary_model$fixed[2,4],
             colour = "blue",
             linetype = "dashed")

```

```{r, fig.height=2.5, fig.width=6}
cog_brms_draws <- cog_brms %>%
  spread_draws(b_adlikePositive, b_Intercept)

mean_intercept = mean(cog_brms_draws$b_Intercept)
sd_control = sd(alldata[alldata$adlike == "Negative", "clinscore"][["clinscore"]]) * 0.1

cog_brms_draws %>%
  dplyr::select(b_adlikePositive, b_Intercept) %>%
  mutate(b_adlikePositive = mean_intercept + b_adlikePositive) %>%
  pivot_longer(cols = c(b_adlikePositive, b_Intercept)) %>%
  group_by(name) %>%
  median_qi(.width = c(.8, .95)) %>%
  ggplot(aes(y = name, x = value, xmin = .lower, xmax = .upper)) +
  geom_rect(xmin = mean_intercept - sd_control,
            xmax = mean_intercept + sd_control,
            ymin = 0, ymax = 3,
            fill = "cornsilk3") +
  geom_pointinterval() +
  geom_vline(xintercept = mean_intercept, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept + sd_control, linetype = "dashed", colour = "blue") +
  geom_vline(xintercept = mean_intercept - sd_control, linetype = "dashed", colour = "blue") +
  scale_y_discrete(labels = c("AD-like positive", "AD-like negative")) +
  labs(x = clinscore_name) +
  theme(axis.title.y = element_blank(),
        axis.text = element_text(size = 15),
        axis.title.x = element_text(size = 15))
```
